Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-10-25 si.tex
first page editing

In this section, we discuss the evaluation of Herzberg-Teller integrals when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933}
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $|e)$ and $|g)$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
The Hamiltonians are quantum harmonic oscillators.
The solutions to the harmonic oscillator Hamiltonian are well known, and their position space wavefunctions are
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N_m H_m(q) \exp(-q^2/2),
\end{equation}
where $N_m = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the $m^{th}$ Hermite Polynomial.\cite{RN230, MorseFeshbach}

For an $n$-quantum excited state vibration $| \tilde{n} \rangle$, and a $m$-quantum ground state vibration $\ket{m}$, the $k^{th}$-order vibrational coupling term is
\begin{equation}
	\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \ \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \ N_n H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k N_m H_m(q) \exp(-\frac{q^2}{2}) \\
		&= N_n N_m \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta).
	\end{split}\label{key}
\end{equation}
The integral corresponds to Franck-Condon factors for $k=0$ and Herzberg-Teller integrals for $k=1$.
The vibrational coupling integrals are written assuming the dipole moment is Taylor expanded about the equilibrium point of the ground state geometry ($q=0$), as done in the main text.
Note that in the limit $\Delta \rightarrow 0$, typical harmonic oscillator selection rules are restored, as the two wells have the same vibrational Hamiltonian and thus share a subspace of orthonormal eigenstates.

\subsection{General Solution to $\langle \tilde{n} | q^k | m \rangle$}
The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, the product $H_n(q-\Delta) q^k H_m(q)$ will also be a polynomial, so that the integral of \autoref{key} will be of the form
\begin{eqnarray}
	\int_{-\infty}^\infty \mathrm{d}q \ P_{nmk}(q) \exp(-q^2 + q\Delta), \\
	P_{nmk}(q) \equiv \sum_\lambda^{n+m+k} a_\lambda q^\lambda
\end{eqnarray}
for some set coefficients $a_\lambda$, to be specified later. 
We can recast the integral into the linear combination of integrals,
\begin{equation}\label{eq:analytic_form}
	\mel{\tilde{n}}{q^k}{m} = N_n N_m \exp(-\frac{\Delta^2}{2}) \sum_\lambda a_\lambda F_\lambda(\Delta),
\end{equation}
where
\begin{equation}\label{integral}
	F_w(\Delta) \equiv \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta).
\end{equation}
\autoref{integral} can be readily evaluated, as we now show. 
 When $w=0$, the integral is evaluated by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		F_0(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4})
	\end{split}
\end{equation}
For the general solution, \autoref{integral} is evaluated via repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\begin{split}
		F_w(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta) \\
		&= \frac{\mathrm{d}}{\mathrm{d}\Delta} F_{w-1}(\Delta)
	\end{split}
\end{equation}
so we can write\footnote{Setting $-x^2 \ = \ \Delta^2/4$ (so that $x \ = \ i \Delta /2$) and using the chain rule ($\frac{\mathrm{d}}{\mathrm{d}x} =  \frac{\mathrm{d}}{\mathrm{d}\Delta}  \frac{\mathrm{d\Delta}}{\mathrm{d}x}$) provides the final result in \autoref{eq:F_solved}.}
\begin{equation}\label{eq:F_solved}
	\begin{split}
		F_w(\Delta) &= \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} F_0(\Delta) \\
		&= \sqrt{\pi} \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} \exp(\frac{\Delta^2}{4}) \\
		&= \sqrt{\pi} \left( -\frac{i}{2} \right)^w \exp(\frac{\Delta^2}{4}) H_w\left(i\frac{\Delta}{2}\right).
	\end{split}
\end{equation}
In the last line we have used the generator definition of Hermite polynomials:  $H_n(x) \equiv (-1)^n e^{x^2} \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$.\cite{MorseFeshbach}

>>>>

In this section, we discuss the evaluation of Herzberg-Teller integrals when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933}
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $|e)$ and $|g)$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
The Hamiltonians are quantum harmonic oscillators.
The harmonic oscillator wavefunctions are well known, and their position space forms are
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N_m H_m(q) \exp(-\frac{q^2}{2}),
\end{equation}
where $N_m = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the $m^\text{th}$ Hermite Polynomial.\cite{RN230, MorseFeshbach}

For an $n$-quantum excited state vibration $| \tilde{n} \rangle$, and a $m$-quantum ground state vibration $\ket{m}$, the $k^\text{th}$ order vibrational coupling term is
\begin{equation}
	\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \ \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \ N_n H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k N_m H_m(q) \exp(-\frac{q^2}{2}) \\
		&= N_n N_m \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ H_n(q-\Delta) q^k H_m(q) \exp(-q^2 + q\Delta).
	\end{split}\label{key}
\end{equation}
The integral corresponds to Franck-Condon factors for $k=0$ and Herzberg-Teller integrals for $k=1$.
The vibrational coupling integrals are written assuming the dipole moment is Taylor expanded about the equilibrium point of the ground state geometry ($q=0$), as done in the main text.
Note that in the limit $\Delta \rightarrow 0$, typical harmonic oscillator selection rules are restored, as the two wells have the same vibrational Hamiltonian and thus share a subspace of orthonormal eigenstates.

\subsection{General Solution to $\langle \tilde{n} | q^k | m \rangle$}
The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, the product $H_n(q-\Delta) q^k H_m(q)$ will also be a polynomial, so that the integral in \autoref{key} will be of the form
\begin{eqnarray}
	\int_{-\infty}^\infty \mathrm{d}q \ P_{nmk}(q) \exp(-q^2 + q\Delta), \\
	P_{nmk}(q) \equiv \sum_\lambda^{n+m+k} a_\lambda q^\lambda
\end{eqnarray}
for some set coefficients $a_\lambda$ to be specified later. 
As a result, we can recast the integral into the linear combination of integrals
\begin{equation}\label{eq:analytic_form}
	\mel{\tilde{n}}{q^k}{m} = N_n N_m \exp(-\frac{\Delta^2}{2}) \sum_\lambda a_\lambda F_\lambda(\Delta),
\end{equation}
where
\begin{equation}\label{integral}
	F_w(\Delta) \equiv \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta).
\end{equation}
\autoref{integral} can be readily evaluated, as we now show. 
 When $w=0$, the integral is evaluated by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		F_0(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4})
	\end{split}
\end{equation}
For the general solution, \autoref{integral} is evaluated via repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\begin{split}
		F_w(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta) \\
		&= \frac{\mathrm{d}}{\mathrm{d}\Delta} F_{w-1}(\Delta)
	\end{split}
\end{equation}
so we can write\footnote{Setting $-x^2 \ = \ \Delta^2/4$ (so that $x \ = \ i \Delta /2$) and using the chain rule ($\frac{\mathrm{d}}{\mathrm{d}x} =  \frac{\mathrm{d}}{\mathrm{d}\Delta}  \frac{\mathrm{d\Delta}}{\mathrm{d}x}$) provides the final result in \autoref{eq:F_solved}.}
\begin{equation}\label{eq:F_solved}
	\begin{split}
		F_w(\Delta) &= \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} F_0(\Delta) \\
		&= \sqrt{\pi} \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} \exp(\frac{\Delta^2}{4}) \\
		&= \sqrt{\pi} \left( -\frac{i}{2} \right)^w \exp(\frac{\Delta^2}{4}) H_w\left(i\frac{\Delta}{2}\right).
	\end{split}
\end{equation}
In the last line we have used the generator definition of Hermite polynomials:  $H_n(x) \equiv (-1)^n e^{x^2} \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$.\cite{MorseFeshbach}
